---
title: "Dealing with second hardest thing in Computer Science"
subtitle: "<br/>How to name things"
author: "Indrajeet Patil"
institute: "esqLABS"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---

layout: false

class: inverse, center, middle

## "There are only two hard things in Computer Science: </br>cache invalidation and naming things."</br>
### - Phil Karlton

---

layout: false

class: inverse, center, middle

# Naming consistency issues

## Case study: *ggplot2*

*ggplot2* is a [plotting framework](https://ggplot2.tidyverse.org/) in R. 

Quite admirably, it anticipates both **British** and **American** English spellings. 

But does it do it consistently?

---

## Function names

Both forms of spellings are supported.

```{r, eval=FALSE}
# works
guide_colorbar(...)

# this works as well
guide_colourbar(...)
```

## Function parameters

Both forms of spellings are supported.

```{r, eval=FALSE}
# works
aes(color = my_variable)

# this works as well
aes(colour = my_variable)
```

---

# Function names *and* parameters 

A user has now come to believe that *ggplot2* supports British and American spellings for both function names and parameters.

And, since they prefer American spellings, they do this:

```{r, eval=FALSE}
guide_colorbar(ticks.color = "black")
```

**That won't work!!**

Because both functions support **only** the British spelling of parameters:

```{r, eval=FALSE}
guide_colourbar(ticks.colour = "black")
guide_colorbar(ticks.colour = "black")
```

This is **inconsistent** and **violates user assumptions** about function and parameter naming schema<sup>*</sup>.

.footnote[

*This can be a nasty source of bug to fix in a function producing complex plots, especially because your first thought is unlikely to be "I must be spelling color incorrectly".

]
