---
title: "Dealing with second hardest thing in Computer Science"
subtitle: "<br/>How to name things"
author: "Indrajeet Patil"
institute: "esqLABS"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

> There are only two hard things in Computer Science: cache invalidation and naming things.
> -- Phil Karlton

---

layout: false

class: inverse, center, middle

# Naming consistency issues

## Case study: *ggplot2*

*ggplot2* is a [plotting framework](https://ggplot2.tidyverse.org/) in R. 

Quite admirably, it anticipates both **British** and **American** English spellings. 

But does it do it consistently?

---

## Function names

Both forms of spellings are supported.

```{r, eval=FALSE}
# works
guide_colorbar(...)

# this works as well
guide_colourbar(...)
```

## Function parameters

Both forms of spellings are supported.

```{r, eval=FALSE}
# works
aes(color = my_variable)

# this works as well
aes(colour = my_variable)
```

---

# Function names *and* parameters 

A user has now come to believe that *ggplot2* supports British and American spellings for both function names and parameters.

And, since they prefer American spellings, they do this:

```{r, eval=FALSE}
guide_colorbar(ticks.color = "black")
```

**That doesn't work!!**

Because both functions support only the British spelling of function parameter:

```{r, eval=FALSE}
guide_colourbar(ticks.colour = "black")
guide_colorbar(ticks.colour = "black")
```

This can be nasty source of bug, especially in a long function. 

It'll also be hard to debug since your first assumption is unlikely to be "I must be spelling color incorrectly".
